Identifying the Tissue of Origin of Extracellular Vesicles Using RNA Expression Signatures

### 2015-09-17 poster/poster.Rnw
fix captions and gene levels in poster figures

<<specific.markers.plot.function,echo=FALSE,results="hide",warning=FALSE>>=
specific.markers.plot <- function(tissue.name,genes) {
    setkey(interesting.gene.reads.wide,"gene_short_name")
    if (missing(genes))
        genes <- as.character(tissue.specific.genes[tissue==tissue.name,gene])
    interesting.tissues <- 
        colnames(interesting.gene.reads.wide)[
           which(apply(apply(interesting.gene.reads.wide[genes,-(1:2),with=FALSE],1,
                             function(x){order(-x)==2}),1,any))+2]
     interesting.tissues <- 
        c("gene_short_name",tissue.name,interesting.tissues)

    tissue.specific.gene.reads <- 
        data.table(melt(interesting.gene.reads.wide[genes,interesting.tissues,with=FALSE],
                        id="gene_short_name",
                        variable.name="tissue",
                        value.name="FPKM"
                        ))
    tissue.specific.gene.reads[,`log_2(FPKM)`:=log2(FPKM+1)]
    return(ggplot(tissue.specific.gene.reads,
                  aes(y=gene_short_name,x=tissue))
           + geom_tile(aes(fill=`log_2(FPKM)`))
           + scale_fill_gradient(limits=c(0,17),
                                 low="white",
                                 high="#002058",
                                 guide=guide_colourbar(title=expression(log[2](FPKM))))
           + scale_x_discrete("", expand = c(0, 0))
           + scale_y_discrete("", expand = c(0, 0))
           + theme_grey(base_size=14)
           + theme(legend.position = "top",
               axis.ticks = element_blank(), 
               plot.margin=unit(c(1,1,1,1),"mm"),
               axis.text.x = element_text(angle = 30, hjust = 1, vjust=1))
           )
}
@

@@

<<uterus.specific.markers.figure,fig.width=8,fig.height=8,echo=FALSE>>=
print(specific.markers.plot("uterus")
      + xlab("Uterus Specific Genes")
      + ylab("Tissue"))
@
}
\captionof{figure}{Expression of uterus specific genes in uterus and
  tissues with the second highest expression for those genes.
  Expression is $\log_2 \mathrm{FPKM}$ and genes are tissue specific
  if they have a tissue specific expression $\tau_g \ge 0.98$. The
  known uterus specific gene PAEP and short RNAs ZNFX1-AS1\_3 (87bp)
  and AL356310.1 (94bp) are highly uterus specific.}

@@

}
\captionof{figure}{Expression of placenta specific genes in placenta
  and tissues with the second highest expression for those genes.
  Expression is $\log_2 \mathrm{FPKM}$ and genes are tissue specific
  if they have a tissue specific expression $\tau_g \ge 0.98$ and
  $\mathrm{FPKM} \ge 1024$ in placenta. The known placenta specific
  PSG genes and CSH genes are all highly placenta specific, and short
  RNAs mir-126, mir-622, and PAPPA-AS1 are fairly placenta specific.}
}

@@

legend("bottomright",       legend=paste0(capwords(a)," ",c("1 Read","10 Reads","1000 Reads", "5000 Reads")),
       col=c("brown","green","#002058", "#E87722"),
       lwd=2)
}
@ 
\captionof{figure}{\ac{ROC} curve of reduced reads in Placenta and
  Uterus samples with the indicated number of total reads. Reads were
  reduced by sampling reads for each sample weighted by the FPKM
  measured for each sample, and run using the same trained SVM
  depicted in Figure \ref{fig:plac_ut_full_roc}. %
  % 
  The placenta \ac{SVM} is fairly accurate at distinguishing between
  samples of placental origin and non-placental origin even when the
  number of reads is very low (10). The uterus \ac{SVM} requires many
  more reads in order to achieve similar accuracy. This is likely due
  to the many fewer uterus specific genes.}
}

>>>>

<<specific.markers.plot.function,echo=FALSE,results="hide",warning=FALSE>>=
specific.markers.plot <- function(tissue.name,genes) {
    setkey(interesting.gene.reads.wide,"gene_short_name")
    if (missing(genes))
        genes <- as.character(tissue.specific.genes[tissue==tissue.name,gene])
    interesting.tissues <- 
        colnames(interesting.gene.reads.wide)[
         apply(interesting.gene.reads.wide[genes,-(1:2),with=FALSE],1,
                             function(x){order(-x)[2]+2})]

    interesting.tissues <- 
        c("gene_short_name",tissue.name,unique(interesting.tissues))

    tissue.specific.gene.reads <- 
        data.table(melt(interesting.gene.reads.wide[genes,interesting.tissues,with=FALSE],
                        id="gene_short_name",
                        variable.name="tissue",
                        value.name="FPKM"
                        ))
    tissue.specific.gene.reads[,`log_2(FPKM)`:=log2(FPKM+1)]
    gene.levels <- 
        unique(as.character(tissue.specific.genes[tissue==tissue.name][order(specificity),gene]))
    tissue.specific.gene.reads[,gene_short_name:=factor(gene_short_name,levels=gene.levels)]
    return(ggplot(tissue.specific.gene.reads,
                  aes(y=gene_short_name,
                          x=tissue))
           + geom_tile(aes(fill=`log_2(FPKM)`))
           + scale_fill_gradient(limits=c(0,17),
                                 low="white",
                                 high="#002058",
                                 guide=guide_colourbar(title=expression(log[2](FPKM))))
           + scale_x_discrete("", expand = c(0, 0))
           + scale_y_discrete("", expand = c(0, 0))
           + theme_grey(base_size=14)
           + theme(legend.position = "top",
                   axis.ticks = element_blank(),
                   panel.grid=element_blank(),
                   panel.border=element_blank(),
                   panel.background=element_blank(),
                   plot.background=element_blank(),
                   plot.margin=unit(c(1,1,1,1),"mm"),
                   axis.text.x = element_text(angle = 30, hjust = 1, vjust=1))
           )
}
@

@@

<<uterus.specific.markers.figure,fig.width=8,fig.height=8,echo=FALSE>>=
print(specific.markers.plot("uterus")
      + xlab("Uterus Specific Genes")
      + ylab("Tissue"))
@
}
\captionof{figure}{Expression of uterus specific genes
  ($\tau_g \ge 0.98$) in uterus and all tissues with the second
  highest expression for any of the uterus specific genes ordered by
  their specificity to uterus. The known uterus specific gene PAEP and
  short RNAs ZNFX1-AS1\_3 (87bp) and AL356310.1 (94bp) are highly
  uterus specific.}

@@

}
\captionof{figure}{Expression of placenta specific genes
  ($\tau_g \ge 0.98$) with high expression ($\mathrm{FPKM} \ge 1024$)
  in placenta and all tissues with the second highest expression for
  any placenta-specific gene ordered by their specificity to placenta.
  The previously known placenta-specific PSG genes and CSH genes are
  highly placenta specific and the short RNAs mir-126, mir-622, and
  PAPPA-AS1 are fairly placenta specific.}
}

@@

legend("bottomright",       legend=paste0(capwords(a)," ",c("1 Read","10 Reads","1000 Reads", "5000 Reads")),
       col=c("brown","green","#002058", "#E87722"),
       lwd=2)
}
@ 
\captionof{figure}{\ac{ROC} curve of reduced reads in Placenta and
  Uterus samples with the indicated number of total reads. Reads were
  reduced by sampling reads for each sample weighted by the FPKM
  measured for each sample, and run using the same trained SVM
  depicted in Figure \ref{fig:plac_ut_full_roc}. %
  % 
  The placenta \ac{SVM} is fairly accurate at distinguishing between
  samples of placental origin and non-placental origin even when the
  number of reads is very low (10). The uterus \ac{SVM} requires many
  more reads in order to achieve similar accuracy. This is likely due
  to the much smaller number of uterus specific genes in comparison to
  placenta.}
}
